Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-06 text/paper/revision_01/paper.tex
K2 reworked section 3 traditionalism

Recently, a lot of evidence based on intuitive judgements and more
theoretical considerations has been discussed in favor of
\mymark{grammaticalism}
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is referred to
as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is assumed to be
similar in effect to the meaning of particle \emph{only} or of the
mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

As this operator can apply at various scope sites, the grammatical
approach predicts that all three readings of \as- and \es-sentences
are available: literal readings arise if no exhaustification operator
is applied; global readings arise if the exhaustification operator
takes sentence-wide scope as in (\ref{Grammar-Global}); and local
readings arise if the exhaustification operator takes scope under the
respective quantifiers as in (\ref{Grammar-Local}).

@@

The grammatical approach, as described so far, does not yet make
falsifiable predictions about the relative salience of our three
candidate readings. In order to derive predictions about reading
preferences, a selection principle is needed that determines which of
the many readings made available by the grammar actually obtain or are
preferred. Proponents of grammaticalism are aware of this and have
suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives.  In the approach of \citet{Magri2011:Another-Argumen},
alternatives may be pruned entirely in the extreme case. In that case,
all occurrences of $\exh$ become vacuous resulting in a literal
reading. However, \citet{Magri2011:Another-Argumen} does not formulate
an explicit theory about which alternatives enter into the computation
of sentence meaning under what circumstances. Therefore, it is not
clear what his approach predicts about the relative salience of the
candidate readings.

Other proponents of grammaticalism assume that insertion of $\exh$ is
optional.  These approaches treat the ambiguity between our candidate
readings as a structural one which resembles syntactic ambiguities. In
order to derive predictions about reading preferences a mechanism of
ambiguity resolution is needed in these approaches. One proposal is
that readings are the more preferred the better they answer the
contextually given question under discussion
\citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}.
Another proposal is to select readings by comparison of logical
strengths, as suggested by the \mymark{strongest meaning hypothesis}
of \citet{DalrympleKanazawa1998:Reciprocal-Expr}
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. For
now, we focus exclusively on this latter selection principle, because
it seems to be the most prominent in the literature and because it
makes the clearest predictions. In the discussion section
(section~\ref{sec:discussion-1}) alternative selection principles will
be reconsidered.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation: fix a sentence with propositional
content $S$ and candidate readings $C(S)$, where $C(S)$ contains $S$
and all readings derivable from inserting $\exh(\cdot)$ at suitable
scope sites;\fn{Strictly speaking, we should compare \emph{parses} of
  sentences with $\exh(\cdot)$ at various scope sites with respect to
  the readings that these parses give rise to
  \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may ignore
  this detail here for ease of exposition.} then define for $X,Y \in
C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Taking stock, under the auxiliary assumptions adopted here,
traditionalism and grammaticalism make different predictions about the
relative salience (and thus preference) of our candidate
readings. Most notably, traditionalism does not predict local readings
for \es-sentences to be available at all while grammaticalism does.
Grammaticalism predicts that all candidate readings are available and,
employing the strongest meaning hypothesis as disambiguation
principle, also comes with a clear prediction about preference:
literal meanings are dispreferred.  All predictions are summarized in
Table~\ref{tab:predictions}.

>>>>

The central idea behind grammaticalist approaches is that quantity
implicatures are derived by application of exhaustivity operators at
varying scope-sites during compositional computation of a sentence's
truth-value, if necessary multiple times
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. For
our purposes, it is enough to assume that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}
%
A grammaticalist core theory predicts that all three readings of \as-
and \es-sentences are in principle available: literal readings arise
if no exhaustification operator is applied; global readings arise if
the exhaustification operator takes sentence-wide scope as in
(\ref{Grammar-Global}); and local readings arise if the
exhaustification operator takes scope under the respective quantifiers
as in (\ref{Grammar-Local}).

@@

It is a topic of active current debate as to how a grammaticalist
approach should be supplemented with a principled \emph{disambiguation
  mechanism} so as to rule out unattested readings obtainable by
embedded exhaustification and to make concrete predictions about
preferences over readings made available. % Grammaticalists therefore
% acknowledge that \emph{some} disambiguation principle must be
% specified, for otherwise grammaticalism would hardly make any
% commitment and would overpredict readings in cases where they are
% unattested, i.e., when scalar items appear in downward-entailing
% contexts \citep[c.f.][for explicit
% discussion]{FoxSpector:Economy-and-Emb}.
Three kinds of approaches have been suggested in the literature
\citep[c.f.][]{ChemlaSingh2014:Remarks-on-the-}. We briefly touch upon
each, but focus on the one that is least controversially applicable to
our experimental case, and that has theoretical significance beyond
the debate about embedded implicatures.

To start with,
\citet{Magri2009:A-Theory-of-Ind,Magri2011:Another-Argumen} opts for
obligatory presence of $\exh(\cdot)$ with variability in the set of relevant
alternatives.  Under this approach, alternatives may be pruned
entirely, so that all occurrences of $\exh(\cdot)$ become vacuous resulting
in a literal reading. Whether alternatives apply or not is a matter of
their relevance. It is not clear what his approach predicts \emph{ex
  ante} about the relative salience of the candidate readings in our
experiment, but we will show in Section~\ref{sec:critical-reflection}
that this approach can be brought to accommodate the data well after
the fact.

Other proponents of grammaticalism assume that insertion of $\exh(\cdot)$ is
optional. Under one proposal for disambiguation, readings are the more
preferred the better they answer the contextually given question under
discussion (QUD)
\citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}. This
is a plausible suggestion, but does not give clear predictions for a
case like our experiment without fixed QUD. Subjects could in
principle accommodate any QUD, so that this approach would not
restrict the space of reading preferences either.

Luckily, the last kind of disambiguation approach does yield a clear
and non-trivial restriction of reading preferences off-the-shelf. It
is to select readings by comparison of logical strengths, as suggested
by the \mymark{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Supplementing
grammaticalism with a strength-based selection criterion has been
suggested, in one form or another, in many contributions to this
paradigm
\citep[e.g.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. At
the same time, the strongest meaning hypothesis has been frequently
called upon for disambiguation purposes in other domains as well,
including reciprocals \citep{DalrympleKanazawa1998:Reciprocal-Expr},
plural predication \citep{Winter2001:Plural-Predicat} or vagueness
\citep{CobrerosEgre2012:Tolerant-Classi}. For these reasons, we focus
from now on this latter type of selection principle. In the discussion
section (section~\ref{sec:discussion-1}) alternative selection
principles will be reconsidered.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation that make non-trivial predications
about preferred readings of \as- and \es-sentences (unlike, say, the
more detailed variants of \cite{FoxSpector:Economy-and-Emb}): fix a
sentence with propositional content $S$ and candidate readings $C(S)$,
where $C(S)$ contains $S$ and all readings derivable from inserting
$\exh(\cdot)$ at suitable scope sites;\fn{Strictly speaking, we should
  compare \emph{parses} of sentences with $\exh(\cdot)$ at various
  scope sites with respect to the readings that these parses give rise
  to \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may
  ignore this detail here for ease of exposition.} then define for
$X,Y \in C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Taking stock, traditionalism and grammaticalism have been introduced
as fairly general core theories that need backup from particular
auxiliary assumptions to make concrete predictions about availability
and readings preferences for our target sentences. Most notably, under
the assumptions made here, traditionalism does not predict local
readings for \es-sentences to be available while grammaticalism does.
Grammaticalism, when supplemented with the the strongest meaning
hypothesis as disambiguation principle, predicts that all candidate
readings are available and that literal readings are dispreferred in
all cases. All predictions are summarized in
Table~\ref{tab:predictions}.
